Data-driven approaches to identifying knowledge gaps in marine invertebrate biodiversity

### 2017-06-01 manuscript.Rmd
add text to manuscript

<!-- Species identification from DNA sequences is poised to revolutionize our understanding of biodiversity, to facilitate environmental monitoring, and to disentangle trophic networks. However, this approach relies on comprehensive databases of accurately identified sequences.  Since its introduction in the early 2000's, DNA barcoding, the sequencing of a common locus across taxa bearing species-specific information, has revitalized taxonomy by facilitating species delineation. More recently, advances in high-throughput sequencing technologies have enabled the parallel sequencing of thousands of species that occur in environmental samples. The approach initially developed to study microbial communities such as the ones from guts or soil, has been modified and extended to be applied to metazoan communities. This approach opens new research venues as it becomes possible to estimate the diversity from species-rich samples that were virtually inaccessible before. This is particularly true in marine communities that include multiple phyla and organisms varying in size by order of magnitude, making identification challenging. However, the utility of the data generated by these projects is enhanced by the possibility of matching the DNA sequences obtained to previously sequenced and correctly identified species. Here, we combined information available in public biodiversity databases to assess the proportion of identified species that have been sequenced in the continental United States. We show that the effort is consistent across the better-studied taxa, with ` average_prop_species_seq`% of the fauna sequenced. Using a DNA barcode dataset of ` n_larvae` planktonic larvae, representing ` n_esus` species, collected in coastal waters of Florida, we find a similar proportion of species that can be assigned to a previously sequenced species ( ` format_output(100 * prop_plankton_with_match)`%).  Our results show that most marine species remain to be sequenced. Achieving this goal will required a concerted effort, and resources. -->

@@

@@

There are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups are well represented in museum collections, the small organisms of the meiofauna are virtually absent: 12 of the 25 phyla represented in the dataset have less than 500 specimens in museums.

@@

The proportion of species that have DNA barcodes available is surprisingly constant across the better studied phyla. About 20% of species known to occur in the Gulf of Mexico and in the Salish Sea have DNA barcode records.  Because geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records)` records analyzed here have no geographic coordinates information associated with them), and many marine invertebrate species are known to be cryptic complexes that are not as widely distributed as previously thought. Therefore specimens identified under the same species but sampled in distant locations, could represent cryptic species. It is thus crucial to ensure that accurate geographic information is provided when DNA barcodes are published, and that additional samples are collected, to capture more accurately their geographic range and the extent of their genetic diversity. This high level of cryptic diversity is reflected in the fact that `r stat_bold$n_spp_muti_bin` species out of `r stat_bold$n_spp` include more than 1 BIN. This number is confounded by some taxonomic identification issues as `r stat_bold$n_shared_bins` BINs are shared across more than one species.

@@

<!-- Paragraph on how sampling effort has changed through time. -->

@@

Recent developments in high-throughput sequencing technologies are giving new applications to these DNA barcode databases [@Leray2016]. Metabarcoding has been developed to sequence in parallel the barcoding gene across the species included in environmental samples such as plankton tows, fish guts, or artificial structure scrappings [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies will soon be used routinely for ecosystem and biodiversity monitoring [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases. Here, we mostly focus on museum specimens because we can validate their identification, and a fraction of these samples (if they have not been preserved in formalin) can be used to obtain DNA from them to complete existing databases of DNA barcodes.

>>>>

<!-- We show that the effort is consistent across the better-studied taxa, with ` average_prop_species_seq`% of the fauna sequenced. Using a DNA barcode dataset of ` n_larvae` planktonic larvae, representing ` n_esus` species, collected in coastal waters of Florida, we find a similar proportion of species that can be assigned to a previously sequenced species ( ` format_output(100 * prop_plankton_with_match)`%).  Our results show that most marine species remain to be sequenced. Achieving this goal will required a concerted effort, and resources. -->

@@

<!-- Species identification from DNA sequences is poised to revolutionize our understanding of biodiversity, to facilitate environmental monitoring, and to disentangle trophic networks. However, this approach relies on comprehensive databases of accurately identified sequences.  -->

@@

There are important discrepancies in our level of knowledge across phyla. While larger, and accessible groups represent the bulk of the records for marine invertebrates in museum collections, the small organisms, including representant of the meiofauna are virtually absent. Of the `r n_idigbio_phyla` phyla occurring in iDigBio, `r records_rare_phyla$n_rare_phyla` are represented by less than 500 records, and make up `r records_rare_phyla$prop_rare_phyla`% of the total number of records.

@@

Since its introduction in the early 2000's, DNA barcoding, the sequencing of a common locus across taxa bearing species-specific information, has revitalized taxonomy by facilitating species delineation [@Meyer2005-plos]. More recently, advances in high-throughput sequencing technologies have enabled the parallel sequencing of thousands of species that occur in environmental samples [@Leray2015]. This approach opens up new research venues it becomes possible to estimate the diversity from species-rich samples that were virtually inaccessible before. This is particularly true in marine communities that include multiple phyla and organisms varying in size by order of magnitude, making identification challenging. However, the utility of the data generated by these projects is enhanced by the possibility of matching the DNA sequences obtained to previously sequenced and correctly identified species [@Leray2015]. With decreasing sequencing costs, and the prospect of automating sample collection, these technologies may soon be used routinely to monitory biodiversity [@Bohan2017]. DNA barcoding provides the possibility to link these newly generated data with historical natural history knowledge about species and their distributions. However, the insights we will gain from these approaches are only going to be as good as the level of completeness and accuracy of the information included in reference databases.

To estimate the proportion of the marine invertebrate diversity for which DNA barcodes are available, we queried BOLD (Barcode of Life Datasystem) for all the species that are listed in the regional lists for the Gulf of Mexico, the Pacific Northwest, as well as all the records for marine invertebrates that occur in the EEZ of the US found in iDigBio. Here, we focused on records based on museum specimens as some of them could potentially be used to complete current DNA barcoding efforts.

The proportion of species that have DNA barcodes available is relatively constant across the better studied phyla at the scale of the EEZ. It varies between 12% for bryozoans to 28% for cnidarians. The proportions for species known to occur in the Gulf of Mexico are usually similar than for the US, while typically a higher fraction of species from the Pacific Northwest are available. Geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records)` BOLD records analyzed here have no geographic coordinates information associated with them), and many marine invertebrate species are known to be cryptic complexes that are not as widely distributed as previously thought. Therefore specimens identified under the same species but sampled in distant locations, could represent cryptic species. It is thus crucial to ensure that accurate geographic information is provided when DNA barcodes are published, and that additional samples are collected, to capture more accurately their geographic range and the extent of their genetic diversity. This high level of cryptic diversity is reflected in the fact that `r stat_bold$n_spp_muti_bin` species out of `r stat_bold$n_spp` include more than 1 BIN. This number is confounded by some taxonomic identification issues as `r stat_bold$n_shared_bins` BINs are shared across more than one species.

```{r proportion-barcoded, fig.cap='Proportion of species known to occur in the US EEZ based on iDigBio records, in the Gulf of Mexico [@Tunnell2009], in the Pacific Northwest[@Kozloff1987], that have publicly available DNA barcodes.'}
knitr::include_graphics("figures/proportion_barcoded.pdf")
```

@@

@@
